---
title: "Lab 04 - La Quinta is Spanish for next to Denny's, Pt. 1"
author: "Jenna Keute"
date: "8 October 2021"
output: html_document
---

### Load packages and data

```
library(tidyverse) 
library(dsbox) 
```

```
states <- read_csv("data/states.csv")
```

### Exercise 1: Determining the dimensions of the Denny's dataset

```{r}
glimpse(dennys)
nrow (dennys)
ncol (dennys)
names (dennys)
```
Each row represents one aspect of an individual Denny's location.
The variables are: "address" "city"  "state"    "zip"   "longitude"  "latitude" 
There are 1643 locations


### Exercise 2: Determining the dimensions of the LaQuinta's dataset

```{r}
glimpse (laquinta)
nrow (laquinta)
ncol (laquinta)
names (laquinta)
```
Each row represents one aspect of an individual Denny's location.
The variables are: "address" "city"  "state"    "zip"   "longitude"  "latitude" 
There are 909 locations

### Exercise 3: Determining how many locations are outside the US

#### Determining how many LaQuinta's locations are outside of the united States

Yes, there are LaQuinta locations outside of the United States
Locations are in Canada, Mexico, New Zealand, Honduras, Turkey, the United States Emirates, Chile, and Colombia

#### Determining how many Denny's locations are outside of the united States

Yes, there are Denny's locations outside of the United States
Locations are in Canada, Mexico, Costa Rica, Honduras, El Salvador, Chili, the United Kingdom, and New Zealand
...

### Exercise 4

While I am unsure of how to implement this, I would start by downloading and viewing the Denny's and LaQuinta's data as a csv file on R.  From there, I would create a "tibbe" on  R that would filter and select the columns based on "country" and state."  However, before I could do that, I would have to create a separate function that would merge all the states names and abbreviations into one file to make it easier when I implement the "filter" command to remove all locations in the US.  Because I know ahead of time that there are so few locations outside the US, I do not think it would be necessary to create another function that would make sorting each country, but that could make organizing the data easier regardless.
...

### Exercise 5

```{r}
dennys %>%
  filter(!(state %in% states$abbreviation))
```
No, there are no locations outside of the US.
...

### Exercise 6

```{r}
dennys %>%
  mutate(country = "United States")
```


...

### Exercise 7

Yes, there are LaQuinta locations outside of the United States
Locations are in Canada, Mexico, New Zealand, Honduras, Turkey, the United States Emirates, Chile, and Colombia

...

### Exercise 8

```{r}
laquinta %>%
  mutate(country = case_when(
    state %in% state.abb     ~ "United States",
    state %in% c("ON", "BC") ~ "Canada",
    state == "ANT"           ~ "Colombia",
    state == "ANT"           ~ "Chili",
    state == "ANT"           ~ "Mexico",
    state == "ANT"           ~ "New Zeland",
    state == "ANT"           ~ "Honduras",
    state == "ANT"           ~ "Turkey",
    state == "ANT"           ~ "United States Emirates"))
```


laquinta <- laquinta %>%
  filter(country == "United States")

### Exercise 9

```{r}
dennys %>%
  count(state) %>%
  inner_join(states, by = c("state" = "abbreviation"))
```


### Exercise 10

```{r}
dennys <- dennys %>%
  mutate(establishment = "Denny's")
laquinta <- laquinta %>%
  mutate(establishment = "La Quinta")

dn_lq <- bind_rows(dennys, laquinta)

ggplot(dn_lq, mapping = aes(x = longitude, y = latitude, color = establishment)) +
  geom_point()
```


